---
output:
  beamer_presentation:
    toc: false
---

# Cultural evolution

Using evolutionary principles to make sense of  
**word origins**,  
**problem solving**, and  
**the growth of Wikipedia articles**.

Pierce Edmiston  
<pedmiston@wisc.edu>  

```{r, include=FALSE}
library(knitr)
opts_chunk$set(
  echo=FALSE,
  message=FALSE,
  warning=FALSE,
  cache=TRUE,
  results="hide"
)

library(magrittr)
library(gridExtra)
library(crotchet)

# Read all R/*.R files in as knitr chunks
dir("R", "*.R", full.names = TRUE) %>% sapply(read_chunk)
```

```{r setup, include=FALSE}
```

# Unifying theme

- **Iteration** is "the repetition of a process or utterance."
- When can iteration be trusted?

# Evolution in the psychology department

Jacobs & Campbell (1961) _J Abnorm Soc Psychol_.  
The perpetuation of an arbitrary tradition through several generations of a laboratory microculture.

```{r jacobs-etal-setup, include=FALSE}
```

# Iterated conformity (Jacobs & Campbell, 1961)

```{r jacobs-etal-fig2-setup}
```

# Iterated conformity (Jacobs & Campbell, 1961)

```{r jacobs-etal-fig2-preds}
```

# Iterated conformity (Jacobs & Campbell, 1961)

```{r jacobs-etal-fig2-data}
```

# Telephone game

```{r telephone-setup, include=FALSE}
```

```{r}
grid.arrange(
  read_image("collect-imitations-gui", "wordsintransition"),
  read_graphviz("definitions", "wordsintransition"),
  nrow = 1
)
```

# Telephone app

```{r}
draw_image("monitor-imitations", "wordsintransition")
```

# Imitations stabilize over generations

```{r telephone-stabilization}
```

# Question types

```{r question-types}
q_true_seed <- read_graphviz("true-seed", "wordsintransition")
q_category_match <- read_graphviz("category-match", "wordsintransition") 
q_specific_match <- read_graphviz("specific-match", "wordsintransition")

grid.arrange(
  q_true_seed,
  q_category_match,
  q_specific_match,
  ncol = 1
)
```

# Guess the seed

```{r guess-the-seed}
```

# Invented words

```{r invented-words, results="asis"}
```

# Category learning

```{r category-learning}
```

# Summary

Unguided repetition of nonverbal imitations makes them more word-like.

- Acoustic form becomes more repeatable and easier to spell.
- Imitations and transcriptions gradually lose resemblance to source.
- As imitations transition into words they become more categorical.

Supports theories of language evolution that value human imitative abilities.

# Problem solving

```{r totems-setup, include=FALSE}
```

```{r strategies}
draw_graphviz("team-structures", "totems")
```

# Technological evolution (Hunt & Gray, 2003)

```{r}
img("hunt_gray/fig4")
```

# Innovation landscape

```{r sample-landscape}
```

# Totems game

```{r totems-game}
draw_image("gameplay", "totems")
```

# Number of inventions

```{r totems-inventory}
```

# Measuring differences in problem solving

- Effectiveness (guesses per invention)
- Redundancy (non-unique guesses)
- Trajectories (unique paths)

# Effectiveness: Guesses per invention

```{r effectiveness}
```

# Redundancy: Non-unique guesses

```{r redundancy}
```

# Trajectories: Exploration of landscape

```{r trajectories}
```

# Prediction

```{r strategies-exp2}
draw_graphviz("team-structures-exp2", "totems")
```

# Is Wikipedia getting better?

> My favorite way of checking this is to "click random article" on 10 articles, and go back and look at them a year ago, 5 years ago, 10 years ago. Every time I have tried, it's unambiguous: Wikipedia is getting better by this test. -- Jimbo Wales

# Wikipedia is hard to measure

- **Size.** Over 5 million articles in English.
- **Expertise.** e.g., [Hurricane Claudette](https://en.wikipedia.org/wiki/Hurricane_Claudette_(2003)).
- **Always changing.** Articles are never considered complete.

# Wikipedia is alive

- [7th most popular website in the world.](http://www.alexa.com/siteinfo/wikipedia.org)
- [6-10 edits per second.](https://tools.wmflabs.org/wmcounter/)
- [700 new articles per day.](https://tools.wmflabs.org/wmcharts/wmchart0002.php)

# Wikipedians

- 140,000 active users (< 30 days).
- Vandalism detection is highly automated.
- New editors do not like getting reverted.
- All editors are protective of their own edits.
- Chance of being reverted doesn't change.

# Wikipedia article editing as an evolutionary strategy

```{r}
grid.arrange(
  read_graphviz("brdc", "wikischolarlib"),
  read_graphviz("edit-tree", "wikischolarlib"),
  nrow = 1,
  widths = c(0.4, 0.6)
)
```

# Experimental evolution (Elena & Lenski, 2003)

```{r}
img("elena_lenski/fitness")
```

# Wikipedia article quality

```{r wikipedia-quality}
```

# Future directions

- Edit quality models (big data!).
- Separate purifying from positive selection.
- Expand to open source software projects.

# Discussion question

What do you like/dislike about the Wikipedia/evolution comparison?

#

Using evolutionary principles to make sense of  
**word origins**,  
**problem solving**, and  
**the growth of Wikipedia articles**.

Pierce Edmiston  
<pedmiston@wisc.edu>  
[github.com/pedmiston/leaning-on-darwin](https://github.com/pedmiston/leaning-on-darwin)

# Proof of principle

```{r mountain-climbers}
```

# Problem solving as hill climbing

```{r mountain-climbers-walks}
```