---
output:
  beamer_presentation:
    toc: false
---

# Cultural evolution

Using evolutionary principles to make sense of  
**word origins**,  
**problem solving**, and  
**the growth of Wikipedia articles**.

Pierce Edmiston  
<pedmiston@wisc.edu>  

```{r, include=FALSE}
library(knitr)
opts_chunk$set(
  echo=FALSE,
  message=FALSE,
  warning=FALSE,
  cache=TRUE,
  cache.path = ".cache/",
  results="hide",
  fig.path = "figs/"
)

library(magrittr)
library(gridExtra)
library(crotchet)

# Read all R/*.R files in as knitr chunks
dir("R", "*.R", full.names = TRUE) %>% sapply(read_chunk)
```

```{r setup, include=FALSE}
```

# Unifying theme

- **Iteration** is "the repetition of a process or utterance."
- When can iteration be trusted?

```{r iteration, engine="dot"}
digraph {
  rankdir=LR
  node[label="" shape="circle" style="filled"]
  a -> b -> c -> d -> e
  e[style=invis]
}
```

# Evolution in the psychology department

Jacobs & Campbell. (1961). _J Abnorm Soc Psychol_.  
The perpetuation of an arbitrary tradition through several generations of a laboratory microculture.

```{r jacobs-etal-setup, include=FALSE}
```

# Iterated conformity (Jacobs & Campbell, 1961)

```{r jacobs-etal-fig2-setup}
```

# Iterated conformity (Jacobs & Campbell, 1961)

```{r jacobs-etal-fig2-preds}
```

# Iterated conformity (Jacobs & Campbell, 1961)

```{r jacobs-etal-fig2-data}
```

# Creating words from iterated vocal imitation

```{r telephone-setup, include=FALSE}
```

```{r}
grid.arrange(
  read_image("collect-imitations-gui", "wordsintransition"),
  read_graphviz("definitions", "wordsintransition"),
  nrow = 1
)
```

# Telephone app

```{r}
draw_image("monitor-imitations", "wordsintransition")
```

# Imitations stabilize over generations

```{r telephone-stabilization}
```

# "Guess the seed" game

```{r guess-the-seed}
draw_image("match-imitations-gui", "wordsintransition")
```

# "Guess the seed" question types

```{r question-types}
q_true_seed <- read_graphviz("true-seed", "wordsintransition")
q_category_match <- read_graphviz("category-match", "wordsintransition") 
q_specific_match <- read_graphviz("specific-match", "wordsintransition")

grid.arrange(
  q_true_seed,
  q_category_match,
  q_specific_match,
  ncol = 1
)
```

# Matching accuracy

```{r matching-accuracy}
```

# Invented words

```{r invented-words, results="asis"}
```

# Category learning

```{r category-learning}
```

# Summary: Word origins

Unguided repetition of nonverbal imitations makes them more word-like.

- Acoustic form becomes more repeatable and easier to spell.
- Imitations and transcriptions gradually lose resemblance to source.
- As imitations transition into words they become more categorical.

```{r iteration, engine="dot"}
```

# Technological evolution (Hunt & Gray, 2003)

```{r totems-setup, include=FALSE}
```

```{r}
img("hunt_gray/fig4")
```

# Strategies

```{r types-of-time, fig.height=4}
```

# Innovation problem solving (Derex & Boyd, 2015)

```{r sample-inventory, fig.align="center"}
draw_graphviz_with_images("inventory", "totems")
```

# Innovation landscape

```{r sample-landscape}
```

# Totems game

```{r totems-game}
draw_image("gameplay", "totems")
```

# Number of inventions

```{r totems-inventory}
```

# Measuring differences in problem solving

- Effectiveness (guesses per invention)
- Redundancy (non-unique guesses)
- Trajectories (unique paths)

# Effectiveness: Guesses per invention

```{r effectiveness}
```

# Redundancy: Non-unique guesses

```{r redundancy}
```

# Trajectories: Exploration of landscape

```{r trajectories}
```

# Summary: Diachronic inheritance

- Diachronic teamwork is the least redundant.
- Synchronic teamwork results in the most attempts.

```{r iteration, engine="dot"}
```

# Total attempts

```{r totems-attempts}
```

# Iteration versus inheritance

```{r iteration-v-inheritance, engine="dot"}
digraph {
  rankdir=LR
  node[label="" shape="circle" style="filled"]
  a1
  a -> b -> c -> d -> e
  a1 -> b1 -> c1 -> d1 -> e1
  e, e1[style=invis]
  b1, c1, d1[style=invis width=0.1]
}
```

```{r kaggle-setup, include=FALSE}
```

# Kaggle competitions

```{r titanic}
```

# Kaggle competition leaderboards

```{r leaderboards}
```

# Submission interval

```{r submission-interval}
```

# Types of strategies

```{r team-types-quartet}
```

# Iteration as a strategy

```{r team-types}
```

# Is Wikipedia getting better?

> My favorite way of checking this is to "click random article" on 10 articles, and go back and look at them a year ago, 5 years ago, 10 years ago. Every time I have tried, it's unambiguous: Wikipedia is getting better by this test. -- Jimbo Wales

# Wikipedia by the numbers

- 5 million articles in English.
- [5th most popular website in the world.](http://www.alexa.com/siteinfo/wikipedia.org)
- [6-10 edits per second.](https://tools.wmflabs.org/wmcounter/)
- [700 new articles per day.](https://tools.wmflabs.org/wmcharts/wmchart0002.php)

# Wikipedians

- 140,000 active users (< 30 days).
- Vandalism detection is highly automated.
- New editors do not like getting reverted.
- All editors are protective of their own edits.
- Chance of being reverted doesn't change.

(For more, start with [Aaron Halfaker](http://www-users.cs.umn.edu/~halfak/)).

# Wikipedia article editing as an evolutionary strategy

```{r}
grid.arrange(
  read_graphviz("brdc", "wikischolarlib"),
  read_graphviz("edit-tree", "wikischolarlib"),
  nrow = 1,
  widths = c(0.4, 0.6)
)
```

# Experimental evolution (Elena & Lenski, 2003)

```{r}
img("elena_lenski/fitness")
```

# Predicting article quality (Warncke-Wang et al., 2015)

```{r wikipedia-quality-classes, fig.width=2.5, fig.align="center"}
draw_image("wikipedia-quality-classes", "wikischolarlib")
```

Authority/Reputation, **Completeness**, Complexity, **Informativeness**, Consistency, Currency, Volatility, Diversity, **NumHeadings**, **ArticleLength**, **NumReferences**, NumWikilinks, HasInfobox, ...

# Monotonic increases in Wikipedia article quality

```{r wikipedia-quality}
```

# Future directions

- Edit quality models (big data!).
- Separate purifying from positive selection.
- Expand to open source software projects.

# Discussion

What do you think of the Wikipedia/evolution comparison?

# Summary

```{r iteration, engine="dot"}
```

- Word origins: Imitations transition to words through repetition
- Problem solving: Diachronic inheritance, Iteration as a strategy
- Wikipedia article editing as an evolutionary strategy

#

Using evolutionary principles to make sense of  
**word origins**,  
**problem solving**, and  
**the growth of Wikipedia articles**.

Pierce Edmiston  
<pedmiston@wisc.edu>  
[github.com/pedmiston/leaning-on-darwin](https://github.com/pedmiston/leaning-on-darwin)

# Proof of principle

```{r mountain-climbers}
```

# Problem solving as hill climbing

```{r mountain-climbers-walks}
```