---
output: beamer_presentation
---

# Cultural evolution

Using evolutionary principles to make sense of  
**word origins**,  
**problem solving**, and  
**the growth of Wikipedia articles**.

Pierce Edmiston  
<pedmiston@wisc.edu>  

```{r, include=FALSE}
library(knitr)
opts_chunk$set(
  echo=FALSE,
  message=FALSE,
  warning=FALSE,
  cache=TRUE,
  results="hide"
)

library(magrittr)
library(gridExtra)
library(crotchet)

# Read all R/*.R files in as knitr chunks
dir("R", "*.R", full.names = TRUE) %>% sapply(read_chunk)
```

```{r setup, include=FALSE}
```

# Unifying theme

- **Iteration** is "the repetition of a process or utterance."
- When can iteration be trusted?

# Evolution in the psychology department

Jacobs & Campbell (1961) _J Abnorm Soc Psychol_.  
The perpetuation of an arbitrary tradition through several generations of a laboratory microculture.

```{r jacobs-etal-setup, include=FALSE}
```

# Iterated conformity (Jacobs & Campbell, 1961)

```{r jacobs-etal-fig2-setup}
```

# Iterated conformity (Jacobs & Campbell, 1961)

```{r jacobs-etal-fig2-preds}
```

# Iterated conformity (Jacobs & Campbell, 1961)

```{r jacobs-etal-fig2-data}
```

# Telephone game

```{r telephone-setup, include=FALSE}
```

```{r}
grid.arrange(
  read_image("collect-imitations-gui", "wordsintransition"),
  read_graphviz("definitions", "wordsintransition"),
  nrow = 1
)
```

# Stabilization of imitations

```{r telephone-stabilization}
```

# Question types

```{r question-types}
q_true_seed <- read_graphviz("true-seed", "wordsintransition")
q_category_match <- read_graphviz("category-match", "wordsintransition") 
q_specific_match <- read_graphviz("specific-match", "wordsintransition")

grid.arrange(
  q_true_seed,
  q_category_match,
  q_specific_match,
  ncol = 1
)
```

# Guess the seed

```{r guess-the-seed}
```

# Category learning

```{r category-learning}
```

# The growth of Wikipedia articles

```{r}
grid.arrange(
  read_graphviz("brdc", "wikischolarlib"),
  read_graphviz("edit-tree", "wikischolarlib"),
  nrow = 1,
  widths = c(0.4, 0.6)
)
```


# 

# Bouba/kiki effect

<p><a href="https://commons.wikimedia.org/wiki/File:Booba-Kiki.svg#/media/File:Booba-Kiki.svg"><img src="https://upload.wikimedia.org/wikipedia/commons/e/e7/Booba-Kiki.svg" alt="Booba-Kiki.svg" height="112" width="220"></a>

- Some words might begin with meanings everyone agrees on.
- Illustrates the general phenomenon of _iconicity_.
- Criticized for being a handpicked example.
- What's the generative process for words like bouba and kiki?